---
title: "Mastering Shiny Notes"
description: |
  Chapter 1: Your first Shiny app
author:
  - name: Alexis Solis Cancino
    url: mailto:alexis.solisc@gmail.com
    # affiliation: ITAM
date: "2021-03-21"
output: 
  distill::distill_article:
    toc: true
    toc_float: true
    code_folding: true
    theme: "resources/theme.css"
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = T,
  warning = F,
  eval = F)

library(shiny)
```


## 1.1 Introduction

Two key components of every Shiny app: the UI (short for user interface) which defines how your app _looks_, and the server function which defines how your app _works_.


## 1.2 Create app directory and file

In this section, Hadley teaches us how to create a super simple app and put it in the `app-01.R` file. The whole code looks like this:

```{r, eval = FALSE}
# NOT RUN

library(shiny)

# User Interface (how it looks)
ui <- fluidPage(
  "Hello, world!"
)

# Server: how it works
server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```


Inspecting the code above, `app-01.R` does four things:

It calls library(shiny) to load the shiny package.

It defines the user interface, the HTML webpage that humans interact with. In this case, it’s a page containing the words “Hello, world!”.

It specifies the behavior of our app by defining a server function. It’s currently empty, so our app doesn’t do anything, but we’ll be back to revisit this shortly.

It executes `shinyApp(ui, server)` to construct and start a Shiny application from UI and server.

RStudio Tip: There are two convenient ways to create a new app in RStudio:

Create a new directory and an app.R file containing a basic app in one step by clicking File | New Project, then selecting New Directory and Shiny Web Application.

If you’ve already created the app.R file, you can quickly add the app boilerplate by typing “shinyapp” and pressing Shift+Tab.

## 1.3 Running and stopping

There are a few ways you can run this app:

Click the Run App (Figure 1.1) button in the document toolbar.

Use a keyboard shortcut: Cmd/Ctrl + Shift + Enter.

If you’re not using RStudio, you can (source())1 the whole document, or call `shiny::runApp()` with the path to the directory containing app.R.

When you run the app, on the R console we find the following code:

```{r}
#> Listening on http://127.0.0.1:3827
```

This tells you the URL where your app can be found: `127.0.0.1` is a standard address that means “this computer” and `3827` is a randomly assigned port number. You can enter that URL into any compatible2 web browser to open another copy of your app.

## 1.4 Adding UI controls



```{r, eval = FALSE}
# NOT RUN
library(shiny)

ui <- fluidPage(
  selectInput("dataset", label = "Dataset", choices = ls("package:datasets")),
  verbatimTextOutput("summary"),
  tableOutput("table")
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

This example uses four new functions:

fluidPage() is a layout function that sets up the basic visual structure of the page. 

selectInput() is an input control that lets the user interact with the app by providing a value. In this case, it’s a select box with the label “Dataset” and lets you choose one of the built-in datasets that come with R.

verbatimTextOutput() and tableOutput() are output controls that tell Shiny where to put rendered output (we’ll get into the how in a moment). verbatimTextOutput() displays code and tableOutput() displays tables.

We only see the input, not the two outputs, because we haven’t yet told Shiny how the input and outputs are related.


## 1.5 Adding behaviour

This is the essence of **reactivity**: outputs automatically react (recalculate) when their inputs change.

```{r, eval = FALSE}
# NOT RUN

server <- function(input, output, session) {
  output$summary <- renderPrint({
    dataset <- get(input$dataset, "package:datasets")
    summary(dataset)
  })
  
  output$table <- renderTable({
    dataset <- get(input$dataset, "package:datasets")
    dataset
  })
}
```


## 1.6 Reducing duplication with reactive expressions

Even in this simple example, we have some code that’s duplicated: the following line is present in both outputs.

```{r}
dataset <- get(input$dataset, "package:datasets")
```


## 1.8 Exercises

1. Create an app that greets the user by name. You don’t know all the functions you need to do this yet, so I’ve included some lines of code below. Think about which lines you’ll use and then copy and paste them into the right place in a Shiny app.

```{r}
# Given code
tableOutput("mortgage")
output$greeting <- renderText({
  paste0("Hello ", input$name)
})
numericInput("age", "How old are you?", value = NA)
textInput("name", "What's your name?")
textOutput("greeting")
output$histogram <- renderPlot({
  hist(rnorm(1000))
}, res = 96)
```

Solution:

```{r}

library(shiny)

ui <- fluidPage(
  # Ask for the name (input)
  textInput("name", "What's your name?"),
  
  # Ask for the age (input)
  numericInput("age", "How old are you?", value = NA),
  
  # Define output 
  textOutput("greeting")
)

server <- function(input, output, session) {
    output$greeting <- renderText({
    paste0("Hello ", input$name)
  })
}

shinyApp(ui, server)
```

2. Suppose your friend wants to design an app that allows the user to set a number (x) between 1 and 50, and displays the result of multiplying this number by 5. This is their first attempt:

```{r, eval = FALSE}
library(shiny)

ui <- fluidPage(
  sliderInput("x", label = "If x is", min = 1, max = 50, value = 30),
  "then x times 5 is",
  textOutput("product")
)

server <- function(input, output, session) {
  output$product <- renderText({
    x * 5
  })
}

shinyApp(ui, server)
```

